### 2017-03-29 bes-write-up-analysis.Rmd
first commit

---
title: "Whales and Isotopes for BES project write-up"
output: html_notebook
date: "`r Sys.Date()`"
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", 
                      fig.width = 8, fig.height = 6)

```
# Meeting Notes 29-March-2017

* 544 samples
    * 1 blue
    * 6 fin
    * 6 minke
    * 3 sei

The data are...

```{r import-data}
# load required libraries
# library(SIBER)

# read in the data
```

# Data summary & visualisation
Summarise the data here...

```{r summary-stats}
# some summary stats

# some basic visualisation

```

# Models

```{r fit-models}
# fit some models

```

# Results
```{r results}
# summarise model results

```

# Some Notebook shortcuts just to remind me

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).