### 2017-11-16 manuscript/revision/response.md
Complete first draft response and edits

# RE: Nature submission 2017-06-07324A

Dear Editors,

Our paper was reviewed at Nature and was not accepted for publication. However, it was suggested that we transfer the paper to Nature Ecology and Evolution, which we are glad to do. We are transferring the reviews with the paper, therefore we include a response to referees as part of this cover letter. We are grateful to the referees for their helpful and constructive comments, and hope that our edits, along with the move from Nature to Nature Ecology and Evolution, will have allayed their concerns about the manuscript.

We respond to their points below. To improve the clarity of this document we have included our responses to the referees comments in italics.

## Important note to the referees 
During the revisions of this manuscript, which included substantial extra analysis, we discovered an error in our original results. Essentially, the baleen samples had been labeled in the reverse direction to what we expected. The values for all the stable isotope measures remain the same but the timeline is reversed (this occurred due to a misunderstanding with the masters student who ran to original mass spec analyses and subsequently finished her masters project and moved away). We regret this error, but these things happen, and we are relieved that we noticed the error before publication. Interestingly, when we added the extra data to the end of our time series, we find our overall results do not change from our previous manuscript. We communicated with Editor in Chief Patrick Goymer about whether it was still appropriate to cascade our reviews with the paper given this error, and he said this was fine as long as we made this very clear in our response letter. This note is to alert the referees that the results graphs will look different to the original version, though our conclusions have not changed qualitatively.

## New analyses
As well as responding to the referee comments, and substantially rewriting sections of the paper, this new version contains several new analyses. In brief:

1. We took further keratin samples from the fronds of the baleen, this has increased our number of samples from 78 to XX, and the time series from six to seven full years of records.
2. We have increased the number of simulations on our models from 100 to 1000, to better explore parameter space. We provide X new figures in the supplementary information exploring these models more thoroughly.

## Original referee comments

## Referee 1
### Key results, Originality and Significance

In my opinion, the paper makes two important points:

First: The authors used a mix between a temporal isoscape (measured POM over time and space to create a changing isoscape) and an agent-based model (AI that models the movement of an individual using parameters such as presence of food, time and temperature in adjoining locations to decide if it moves or not). The combination was used to retrace the movement of a blue whale and compare it to baleen isotope records. In my opinion, this is a novel approach that provides more confidence in attributing isotopic signals to place/time, and thus helps reconstruct migration of the individual. However, I feel that it is not a solve-all approach, as just as with isoscapes, you still need to get those initial measurements of time/space isotopes. Nevertheless, it is a very interesting technique.

Second: the baleen isotope records themselves are interesting as there are very few blue whale baleen records published (this is the second one ever to my knowledge, and the first was also published this year). The isotopes also appear to show a mixture of behaviours, such as migration AND permanent residence in a single individual. Such mix of behaviour is rare in the literature, and still debated. Further, the residence appears consistent with pregnancy/lactation, which is even more uncommon in whale baleen literature as far as I know. For these reasons, I think the isotope records will be very interesting to other researchers in the field.

> We thank the referee for their positive comments about our paper.

### Data and Methodology: Validity

I am satisfied with the SIA methods and results, but I have some slight comments. While I have no experience with modelling, I noted that there was no mention of isotopic fractionation in the model. As far as I know, blue whale feed on zooplankton and small fish, not directly on phytoplankton. If I am correct, the comparison of the baleen records to the modelled isotopes may be incorrect, as it should be affected by trophic shifts. I suppose it depends on whether the model was used for a direct pattern comparison, or if the values modelled and the baleen values were compared to each other. 
> The referee is correct that blue whales feed predominantly on krill (small crustaceans), not directly on phytoplankton. However, the trophic shift from phytoplankton to crustacean predators is very small

* CLIVE how are we dealing with this???

Comparable literature in other species (seals, humans) appear to confirm the pregnancy interpretation, although I'm less convinced by the lactation signal as it seems too delayed in time. This could possibly be checked using hormones, although that is also a very novel field. See the following literature on baleen hormones: Hunt 2016 “Longitudinal progesterone profiles in baleen from female North Atlantic right whales (Eubalaena glacialis) match known calving history”; Hunt 2014 “Baleen hormones: a novel tool for retrospective assessment of stress and reproduction in bowhead whales (Balaena mysticetus)”. These are older samples (20 yrs or so) but not as old as this whale so may be loss of hormone signal over time.
I strongly believe this will not prohibit publication, as the results are still very interesting. It is very possible that I am misreading the delay in lactation signal.

> Measuring hormone levels in the baleen would be a great addition to this work. We have colleagues in the US who are planning to extract hormones from this baleen. Unfortunately these results have been delayed now for over a year due to US permitting issues (thanks Trump!) and the CITES status of blue whales. But we agree that this would be really interesting, if the hormone signals remain over time.

> In terms of the lactation signal, the reviewer makes a good point, and the changes to our results in this version also fail to demonstrate a convincing signal of lactation. We have therefore removed the mention of the lactation signal from the manuscript. We still see evidence of pregnancy and have retained this interpretation.

### Appropriate use of statistics:

The statistical analyses seem sound and appropriate, although the use of Fourier analysis to calculate periodicity and growth rate in baleen can be difficult with the typically low number of samples per cycle and replication. However, since most of the literature is based on “eyeballing” cycles and visual estimates, the Fourier analysis is a good way to confirm the visual estimates.

There are no error bars or replication of samples, but this is typical of stable isotopes.

> Thank you

### Suggested improvements/ References:

Some basic stable isotope knowledge, as well as blue whale feeding ecology should probably be mentioned (ie krill/fish feeding, no fasting) or trophic influence on isotopic signal compared to model POM values. Adding sentence on feeding ecology at beginning, and explaining in model why POM is a good proxy for zooplankton should be enough. For example: Line 65: a reference such as West et al, 2006 “Stable isotopes as one of Nature’s ecological records” relevant to SIA basics. Line 82: reference again with basic SIA shifts in marine food webs? Possibly McMahon et al, 2015 “Millennial Scale plankton regime shifts in the subtropical North Pacific Ocean”.

> We have added these as suggested.

* In line XX we specifically mention that blue whales feed on krill ("Blue whales feed mainly on krill (Euphausiacea), small crustaceans that feed on phytoplankton."). 
* In lines XX we mention the trophic influence on isotopic signatures [NEED to DO this]
* We have added the West et al reference at lines XX and XX
* We have added an explanation of d13C and d15N at line XX: "Note that carbon and nitrogen isotopes are expressed as d13C and d15N values, i.e. relative to a standard and as ratios of the rare (13C, 15N) to common (12C, 14N) isotope"
* We have added the McMahon reference at line XX

The reported baleen growth rates fit the literature, but I believe it may be worth mentioning in Line 61 that baleen gets worn down over time, and that the records are not full life spans.

> We have clarified this in the text at line XXX as follows: "It is ideal for stable isotope studies because keratin grows continuously through an individual's life (although it is worn away at the tips over time, so rarely records an individual's entire lifespan), and once laid down it is metabolically inert."

According to the model, d15N is relatively higher in the arctic during summer than the Atlantic in winter. In the profile, the winter values in phase one are still comparatively higher than the summer values. Would you be able to explain? I am mostly curious, but it may reflect difficulties with reading the model from a figure.

> Explain. I'm not sure what is going on here - Clive?

While the d13C increase fits the sea lion theory, the drop in d13C seen after/at birth happens too quick (late 1889) in the whale to fit the d15N rise consistent with lactation (Almost 6 months late according to timeline, past mid 1890). Human analyses in hair showed a decrease in d15N (1%o) during pregnancy, which can be seen in the baleen if we take the d13C drop as birth time and early 1889 as pregnancy time (see Fuller et al 2004 “Nitrogen Balance and 15N, why you're not what you eat during pregnancy”). While it's not a study conducted in marine mammals, it's a further indication that the birth time may indeed be in late 1889, or at least that the lactation peak is very late. Could indicate a slowing down in baleen growth, or something else? Could be worth checking with hormonal analyses (see above).

> As mentioned above, we are unable to complete hormonal analyses at this time (and are unsure if the signal will remain for over 100 years). However, we agree with the reviewers concerns and have removed mention of lactation signal from the manuscript, particularly after our reanalysis removed the strength of any signal.

Another possibility lies in the location of baleen analyses. Since the methods are not clear on that point, can you confirm that the samples were taken past the gum line? Dried skin/gum could have affected the isotopic signals if samples were taken outside of the erupted zone. Could it be why the isotopic signals are so different in phase two? Busquet Vass paper shows no difference in 13C between skin and baleen so unlikely, but worth confirming.

> We sampled starting at 0.5cm from the proximal end of the baleen, which is likely to cross the gum line. However, the baleen plate was cleaned thoroughly before analysis, and even prior to this no trace of skin or gum could be seen on the plates (this material was carefully cleaned back in the 1800s when the baleen was first stored). We therefore do not believe this reflects skin or gum influencing the isotopic signals. We had mentioned cleaning the plate in the methods previously, but have now explicitly mentioned that this cleaning removes skin/gum (and that this can influence stable isotope readings) at line XX: "The baleen plate was cleaned with ethanol to remove surface contaminants such as skin/gum or other lipids that can influence isotopic signals."

The final sample interpretation of a return to cyclicity is very open to discussion, as a single point is too weak to be a guarantee. I will concede that the isotopes are consistent with past migratory values, but may be worth softening the comment (line 109-110)?

> This is a fair point. We have now edited this whole section because of the reverse of the timescale of the analysis and so have removed this comment.

When describing the different behaviours, such as whales staying in high latitudes instead of migrating, also talk about the ones staying at low lat (isotope blue whale ref. Busquet-Vass) since it is more relevant to your own isotopic interpretation (line 127).

> We have added this reference.

Typos etc:

1. Abstract
Line 20: Should not have abbreviation. Move to line 58 where it is introduced next.

> Fixed

2. Article main text
Line 56: “)” extraneous bracket

> Fixed

3. Supplementary Data
Page 2, line 1: “)” extraneous bracket after “behavioural phase 2”

> Fixed

### Clarity and Context:

The manuscript was clear, but lacked some information on basic isotope literature, and blue whale ecology. While the ecology side is less important, some information on how isotope analyses work could be beneficial to the greater public. I understand that the limited space available prevents a thorough explanation of SIA, but there is no explanation of the general differences between d13C and d15N (geographical versus trophic information). I believe a quick sentence describing the two tracers in line 57 (for example) could help.

> We have added a brief comment on what blue whales eat on line XXX: "Blue whales feed mainly on krill (Euphausiacea), small crustaceans that feed on phytoplankton".

> We have also added some basic information on stable isotopes, but have tried to avoid increasing the word count too much. We clarify the units used as this often trips up new users. line XXX "We then used continuous-flow isotope ratio mass spectrometry to determine the isotopic compositions of carbon and nitrogen in the keratin. These isotopes are expressed as d13C and d15N values, i.e. relative to a standard and as ratios of the rare (13C and 15N) to common (12C and 14N) isotope (West et al 2006).
Carbon and nitrogen isotopes are commonly reported together as they provide complementary information; isotopes provide information on the geographical locations of an individual, whereas nitrogen isotopes provide trophic and physiological information (West et al 2006)."

The abstract was clear and direct, providing a good summary of the paper's approach and conclusions.
The conclusion is clear and justified, although the claim of first quantitative evidence of movement behaviours in blue whale seems unlikely given the slightly older paper from Busquets-Vass also reporting blue whale baleen isotopes.

> We have removed this comment from the manuscript, and discuss the novelty of our work more in the response to referee 2 below.

## Referee 2:

> If we are interpreting Referee 2's comments correctly (see below), we think they mostly revolve around issues about the novelty of the study, and overstatement of results based on one individual. We completely see the points the referee is making. Some of this was related to the difficulties of writing in the shorthand Nature style - we have clearly been a little too casual in some of our wording. In other cases we haven't been clear that the selling point of this method is that it can easily be applied to hundreds of individuals at relatively low cost. Here we test the methods on one individual, but the real value in the method is its broader applicability. We have tried to clarify this where possible and have also toned down/removed/clarified a lot of our previous statements about the novelty of the study and interpretations of the results. We explain our changes in more detail below, and apologise for any offense caused by our apparent disregard for previous excellent work in this area.

I appreciated the opportunity to review the submitted manuscript “Reconstructing the last known movements of one of Nature's giants”.

The authors use stable isotope analysis (SIA) of baleen of a single blue whale, stranded in 1891, to reconstruct a time series of ~6 years. This isotopic reconstruction represents the integrated isotopic ‘signal’ the whale encountered as it moved, allowing reconstruction of past movements. The authors paired these isotopic data with an ‘isoscape’ of values in the North Atlantic, as well as a model that simulated possible whale migration patterns over the isoscape. By combining these two approaches the authors suggest north-south movement for several years, followed by residency in subtropical waters. The authors claim that the similarity of these movements with modern blue whales suggest that movement patterns of blue whales in the North Atlantic have been similar for hundreds of years, despite anthropogenic pressures (e.g., whaling).

The paper is interesting and as the authors state, provides potentially valuable information on (one) historical blue whale movements in the North Atlantic Ocean. However, I have substantial concerns regarding the originality/novelty, assumptions, and conclusions of the study, enough so that I cannot recommend its publication in Nature.

The Abstract and text explicitly or implicitly state the novelty of the following: “first quantitative multi-year records of historical migrations in individual blue whales”, similarity of blue whale movements over hundreds of years, use of archived museum specimens, novel agent based simulation models.

The study does not provide "multi-year records of...individual blue whales", but of one blue whale. The generalization of the results from one individual to the entire population is a central issue to the current paper.

> This is a fair point, and partially the result of some lax wording on our part. We have toned down our claims of novelty throughout and removed this sentence from the abstract.

There are also issues with the movement interpretations for this one whale; the most important is that the authors make strong conclusions re: movement when the measured SIA values match those predicted by their model, but then employ a physiological argument for 'phase 2' (pregnancy and lactation) when there is a mismatch between measured data and the model. Of course, this approach makes the assumption that physiology can override the environmental signal, which calls into question any inferred movement patterns throughout the time series.

> The referee is correct in that ???

A main conclusion of the paper, that blue whales then migrated as they do now, simply cannot be made from one whale. I appreciate the difficulty of obtaining specimens like the one here, which makes larger sample sizes difficult to obtain. That limitation does make n = 1 studies reasonable, worthwhile, and interesting; it is my opinion that the results here, both for insights into the movements of this whale and for the methodology employed (combining a movement/isoscape model with measured values), should be published. However, n = 1 studies cannot be used to generalize the movements of an entire wild, historical population.

> This is a fair point again. As stated above this is partially an issue of our wording and attempts to write this in a Nature style. We have toned down these comments throughout, and focus instead on promoting the methodological approach we have used, which we do believe is novel.

This paper is also not the first to use historical data to reconstruct past movements of blue whales; see for example Reeves et al (2004) that addresses the same question with whale landings data.

As the authors cite, the reconstruction of whale movements using baleen is well established in the literature; see papers by Best, Schell, Hobson, and others for similar analyses on Bowhead, Right, and other baleen whales.

> Yes we are aware of these other papers, and cite many of them already. We agree that our wording in the original manuscript was not very clear and overly stated the novelty of our findings. We have tried to tone this down throughout, and provide more balanced coverage of previous work.

The penultimate sentence of the abstract and final sentence of the paper suggest that using archived museum specimens is a novel concept in marine ecology. The use of archived museum specimens is not new or overlooked; two immediate examples (of many) include Vo et al (2010) using century-old seabirds and Yeakel et al (2009) on the famous lions of Tsavo. There are papers written that point to the value of using archived collections for data (e.g. Lister 2011), and the approach is common enough that there is a methods paper written on the preparation of archived specimens for SIA (Edwards et al 2002).

> Again we realise that our wording was lax in places and overstated the novelty of this approach. This was not our intention - our intention was to encourage more ecologists to consider these archives of data in their work. We are often surprised by how under-utilised our collections are for this kind of work (even in the NHM) so wanted to highlight this. We have toned down our claims of novelty and have added more references to clarify that this is an established field of research.

The “novel” movement model using SIA is indeed an improvement over many past studies using SIA to reconstruct animal movements. However, there are some studies have employed similar methods (e.g. Flockhart et al 2013 in Monarch butterflies).

> OK ???? HAVE THEY???

In short, the authors applied a common tool in an established way to a single historical specimen of a rare species (though their agent-based movement model is a significant improvement over similar studies). Their model matched their measurements sometimes, but not always, leading to a movement-based interpretation for one period and a physiology-based interpretation (pregnancy and lactation) for the second period.

Perhaps most importantly, if the results and conclusions are considered completely correct, the authors still provide the movements of only a single whale. The conclusion that historical blue whale movements were the same as today's based on a single individual is simply not appropriate.

> We have removed this from the text and have toned down interpretations that relate to just a single individual.

By not overextending their results and by addressing a few issues, I do think this could be an interesting and valuable addition to the isotope and marine mammal literature.

> We hope our edits have addressed these issues.

See below for additional, specific comments.

Abstract: the inference of pregnancy and lactation seems more questionable than as it is stated here.

> We have removed the reference to lactation, and reworded to make it clear that the pregnancy conclusion is more tenuous than it appeared in the original abstract.

As mentioned above, the Abstract overextends the results (e.g., "migrations in blue whales" shouldnt be plural; "hundreds of years" which is not the timeline they present, etc).

> We have reworded the abstract and removed this inaccurate comment.

Intro: migratory species may be affected by climate change by modifying movements, but aren't necessarily 'vulnerable'. Some would argue that migratory species are the least vulnerable to climate change.

> We have reworded this to "Such species may also be particularly affected by climate change".

lines 51-52 criticize approaches that are limited to only a few individuals, but this paper provides information on only one individual.

> Fair point. We were trying to clarify that our approach is more scalable, but agree that this was poorly worded. Instead we now state on line XXX: "These methods can be applied to large numbers of individuals, but as proof of concept, here we apply this method to baleen from the Natural History Museum (NHM) London's blue whale `Hope'."

Fig 1 is not very clear. There is no clear indication of the whales movement, and no isotope data on the Figure. I understand the isoscape is in the SM but this figure could be improved to provide better movement and/or isotope data.

> We are a little confused by this comment. I assume the referee meant Figure 2? Figure 1 is a trace of d13C and d15N, and quite clearly contains isotope data. 
FIX!!!

The movement simulations in Fig 2 are also confusing. The SM clarifies but this Figure could show more. For example, the data matches the simulations...of what? The reader cannot tell if the differences in simulations are due to isotopic variability or different movement scenarios. Perhaps the 4 movement scenarios shown in SM could be 4 different colors here. In that case, when the measured data match a given slope, peak, or trough, the reader can associate it with a geographic region.

> Again, we assume the referee is referring to Fig 3 not Fig 2. This figure has changed now we have changed our movement models and added more data. We have clarified what the figure is showing in the legend as follows: :
FIX!!!

lines 112-122. The conclusion of an isotopic reflection of pregnancy and lactation is stated rather strongly here given the evidence (the authors describe the inference as 'tentative' in the SM).

> We have rephrased this paragraph in response to the new data and analyses we have added. We have also toned down the claims of pregnancy and removed those about lactation.

128: "resident" to where?

> We have removed this comment from the text.

137: this point made extensively above....describes the overall "spatial ecology of blue whales in the North Atlantic" based on isotopic inferences over 6 years in a single whale.

> We have removed this statement and toned down claims based on this one whale throughout.

139: Has it been demonstrated that whaling can actually modify movement patterns? Should be stated if so.

> Not per se, though we see changes in pre and post whaling distributions, as seen in Reeves et al 2004. We have modified this statement.

140-141: OK...but it is movement behavior in a single blue whale, and this has been done in larger sample sets of other baleen whales.

> We have removed this statement and toned down claims based on this one whale throughout.

147-149. This sentence makes it seem like hardly anyone is using museum specimens for SIA or other chemical analyses, which simply isn't the case.

> We have rephrased this to make it clear we are highlighting something others have also highlighted in the past.

> We hope we have responded to all these comments appropriately. Please let us know if you require any further information,

> Clive Trueman, on behalf of the authors.