### 2018-11-07 code/pacific-figure-andrew-temp/plot-figure-x.Rmd
added a temp folder with my code and data for the pacific figure
if we are happy with this, i can move the data to the data folder and
move this up a folder and adjust the paths accordingly. For now ive
just dumped it in here as a block. One question mark is over the
workflow, which doesnt show where the files test2A.csv etc… come from,
which i think is from some of Clive’s original code. I can check this
too.

---
title: "Plot-figure-X"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
---
title: "generates Figure X"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Setup

Load required libraries.

```{r setup}

library(ggplot2)
library(dplyr)

```

## Read in the data

This data comes is derived from processing the Pacific whale movement models. These are the best fit simulations in terms of $R^2$ values obtained from linear regression of the normalised model predictions on the normalised empirical data.

```{r import-empirical-data}

# data in local csv file
pacific_data <- read.csv("pacific-blue-whale-isotope-data.csv", 
                         header = TRUE, stringsAsFactors = FALSE)

# convert ID to factor
# pacific_data$whale_ID <- factor(pacific_data$whale_ID)

# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
# set the day no based on estimated growth rate given in Busquetts-Vass
GRx<-15.5
GRA<-13.5
GRB<-14.8
GRC<-17.5

# bundle into an ordered vector
growth_rates <- c(GRA, GRB, GRC, GRx, GRx, GRx)

# create an empty vector in the empirical data.frame to record day number (time)
pacific_data$day_no <- 0
pacific_data$day_no_f <- 0 # flipped time vecrot

# loop over each individual and calculate their spefic time vector 
# based on their growth rates
for (i in 1:length(growth_rates)){

  # find indices of focal whale individual
  idx <- pacific_data$whale_ID == LETTERS[i]

  # conver their sampled baleen cm to time via growth rate
  pacific_data$day_no[idx] <- as.integer(pacific_data$sample_cm[idx] * 
                                           growth_rates[i]) + 1

  # flip time into day_no_f vector
  pacific_data$day_no_f[idx] <- -1 * (pacific_data$day_no[idx] - 
                                        max(pacific_data$day_no[idx]))

}

# 
# # baleen$day.no<-baleen$cm*GRx
# 
# colz<-colorRampPalette(c("brown", "cadet blue", "pink"))(12)
# colz<-seq(1:6)
# A<-baleen[baleen$Baleen_code=="A",]
# B<-baleen[baleen$Baleen_code =="B",]
# C<-baleen[baleen$Baleen_code =="C",]
# D<-baleen[baleen$Baleen_code =="D",]
# E<-baleen[baleen$Baleen_code =="E",]
# Fw<-baleen[baleen$Baleen_code =="F",]
# 
# A$day.no<-as.integer(A$cm*GRA)+1
# B$day.no<-as.integer(B$cm*GRB)+1
# C$day.no<-as.integer(C$cm*GRC)+1
# D$day.no<-as.integer(D$cm*GRx)+1
# E$day.no<-as.integer(E$cm*GRx)+1
# Fw$day.no<-as.integer(Fw$cm*GRx)+1
# 
# # fix so that day numbers start at the oldest end (i.e. time moves forwards as in simulated baleen)
# A$day.no.F<-((A$day.no)-max(A$day.no, na.rm=TRUE))*-1
# B$day.no.F<-((B$day.no)-max(B$day.no, na.rm=TRUE))*-1
# C$day.no.F<-((C$day.no)-max(C$day.no, na.rm=TRUE))*-1
# D$day.no.F<-((D$day.no)-max(D$day.no, na.rm=TRUE))*-1
# E$day.no.F<-((E$day.no)-max(E$day.no, na.rm=TRUE))*-1
# Fw$day.no.F<-((Fw$day.no)-max(Fw$day.no, na.rm=TRUE))*-1

```

## Quick plot of empirical data

```{r quick-plot}

p1 <- ggplot(pacific_data, aes(x = day_no_f, y = d13C)) + 
  facet_wrap(~whale_ID) + 
  geom_line(col = "red") + 
  geom_point(col = "red", size = 1)

print(p1)

```

## Import simulated data

```{r}
# Read in the corresponding best fit simulated data

best.start.All<-c(7, 1, 6, 7, 1, 5)
end.All<-c(41, 67, 67, 55, 71, 58)
best.sim.All<-c(197, 14, 121, 48, 41, 48)

# they are in csv files named test2A, test2B etc...
read_in_these <- paste0("test2", LETTERS[1:6], ".csv")

# read in and bind into a single data.frame
sim_data <- list()

# loop over csv files, import and subset for the "best sim"
for (i in 1:length(read_in_these)){

  tmp <- read.csv(read_in_these[i], 
                                 header = TRUE, 
                                 stringsAsFactors = FALSE)

  # add the appropriate whale_ID letter
  tmp$whale_ID <- LETTERS[i]

  # subset by the index for the best sim in each case
  # this is hard-coded from above
  tmp2 <- subset(tmp, tmp$Rep == best.sim.All[i])

  # further select a subset window of this time series
  tmp3 <- tmp2[best.start.All[i]:(end.All[i]+(best.start.All[i]-1)),]

  # add the time vectors from the empirical data that should match
  tmp3$day_no <- pacific_data$day_no[pacific_data$whale_ID == LETTERS[i]]
  tmp3$day_no_f <- pacific_data$day_no_f[pacific_data$whale_ID == LETTERS[i]]

  print(unique(tmp3$whale_ID))

  # write this subset of data into the list sim_data and for ad hoc reasons
  # remove the NAs from the end
  sim_data[[i]] <- tmp3[!is.na(tmp3$whale_ID),]

  # remove tmp for next iteration of loop
  rm(tmp, tmp2, tmp3)
}

# cant use lapply easily owing to the required subsetting
# sim_data <- dplyr::bind_rows(lapply(read_in_these, 
#                                     read.csv,
#                                     header = T,
#                                     stringsAsFactors = FALSE))

# convert list to single data.frame
sim_data_df <- bind_rows(sim_data)

# coerce whale_ID to factor
sim_data_df$whale_ID <- factor(sim_data_df$whale_ID)

# remove no longer required objects
rm(sim_data)

```

## Quick plot simulated data

```{r quick-simulated}

p2 <- ggplot(sim_data_df, aes(x = day_no, y = TL2)) + 
  facet_wrap(~whale_ID) + 
  geom_line()

print(p2)

```

## plot both together

```{r plot-both-together}

# merge them 
merged_whales <- pacific_data %>% 
  left_join(sim_data_df, by = c("whale_ID", "day_no", "day_no_f"))

merged_whales <- merged_whales %>% group_by(whale_ID) %>% 
  mutate(
  z.d13C.obs = scale(d13C.x),
  z.d13C.sim = scale(TL2)
)

p3 <-  ggplot(merged_whales, aes(x = day_no_f, y = scale(d13C.x))) + 
  facet_wrap(~whale_ID) + 
  geom_line(col = "red") + 
  geom_point(col = "red", size = 1) + 
  geom_line(aes(x = day_no, y = scale(TL2)), col = "black") + 
  xlab("Day number") + 
  ylab("z-scored d13C")

print(p3)

# plot the data z-scored by group

p4 <-  ggplot(merged_whales, aes(x = day_no_f, y = z.d13C.obs)) + 
  facet_wrap(~whale_ID) + 
  geom_line(col = "red") + 
  geom_point(col = "red", size = 1) + 
  geom_line(aes(x = day_no, y = z.d13C.sim), col = "black") + 
  xlab("Day number") + 
  ylab("z-scored d13C")

print(p4)

ggsave("pacific-fits.pdf", plot = p4)
```